---
title: "Project 1"
author: "Guy Merlin Fomen"
date: "27 avril 2019"
output: html_document
---
```{r, echo=FALSE}
library(dplyr)
library(lattice)
```

#I. Loading and Processing the data 
### Code
```{r, echo=TRUE}
activity <- read.csv("activity.csv")
```


#II. Mean Total Number of steps Per day
###1. The total number of steps taken each day
```{r, echo=TRUE}
total_steps<- activity %>% group_by(date) %>% summarize(total_day = sum(steps, na.rm = TRUE))
```


###2. Histogram
```{r, echo=TRUE}
hist(total_steps$total_day, main = " Histogram of the total number of steps each day", xlab = "Total Number of steps", col = " lightblue")
```


###3. Mean and Median
* Mean 
```{r, echo=TRUE}
m1 <- mean(total_steps$total_day)
m1
```

* Median
```{r, echo=TRUE}
Med1 <- median(total_steps$total_day)
Med1
```


#III. Average daily Activity patern

###1. Time serie Plot
```{r, echo=TRUE}
avg_steps<- activity %>% group_by(interval) %>% summarize(avg_day = mean(steps, na.rm = TRUE))

plot(avg_steps$interval,avg_steps$avg_day, type = "l",col="blue",xlab="Time", ylab=" Average", main=" AVerage Steps by days")
u <- which(avg_steps$avg_day==max(avg_steps$avg_day,na.rm = TRUE ))

points(x=avg_steps$interval[u],y=max(avg_steps$avg_day,na.rm=TRUE),type="p",pch=19,col="green")
points(x=avg_steps$interval[u],y=0,type="p",pch=19,col="green")
points(x=0,y=max(avg_steps$avg_day,na.rm=TRUE),type="p",pch=19,col="green")

segments(0,max(avg_steps$avg_day,na.rm = TRUE),avg_steps$interval[u],max(avg_steps$avg_day,na.rm = TRUE),col = "green", lty = 5)
segments(avg_steps$interval[u],0,avg_steps$interval[u],max(avg_steps$avg_day,na.rm = TRUE), col="green", lty = 5)
grid(30,30)
avg_steps$interval[u]
```


###2. Interval for the maximum number of daily steps
The 5-minutes interval on average is[830,835]

#IV. Imputing Data
###1. Total number of row with NA
```{r, echo=TRUE}
totNa <- length(which(is.na(activity$steps)))
totNa
```

###2. Strategy to fill out the NA
We will replace the NA by the mean of over all steps and create the new dataset
```{r, echo=TRUE}
activity$steps[which(is.na(activity$steps))] <- rep(round(mean(activity$steps, na.rm = TRUE),0),totNa)
```

###3. Histogram
```{r, echo=TRUE}
total_st<- activity %>% group_by(date) %>% summarize(tot_day = sum(steps, na.rm = TRUE))
hist(total_st$tot_day, main = " New Histogram of the total number of steps each day", xlab = "Total Number of steps", col = " lightblue")
```

* Mean 
```{r, echo=TRUE}
m2 <- mean(total_st$tot_day)
m2
```

* Median
```{r, echo=TRUE}
Med2 <- median(total_st$tot_day)
Med2
```


The Mean and the Median are now different from the previous one above.

# The Impact of the Imputs
```{r, echo=TRUE}
abs(m1-m2) # for the means
abs(Med1-Med2) # for the Median
```


#V. Activities differences between Weekdays and Weekends
###1. Weekday and Weekend factors
```{r, echo=TRUE}
activity$week <- ifelse(weekdays(as.Date(levels(activity$date))[activity$date]) %in% c("samedi", "dimanche"), "weekend", "weekday")
```

###2. Time serie plot
```{r, echo=TRUE}
avg_steps2<- activity %>% group_by(interval,week) %>% summarize(avg_day = mean(steps, na.rm = TRUE))
xyplot(avg_steps2$avg_day~avg_steps2$interval|avg_steps2$week, avg_steps2,layout=c(1,2), type="l", xlab = "Time", ylab="Average")

```

